{\helveticaitalic{Supplementary Material}}: \\ \helvetica{Xolotl: An Intuitive and Approachable Neuron \& Network Simulator in MATLAB}

### 2018-07-31 paper/paper.tex
Add to supplementary information, fix some spelling

In this example, we demonstrate how to simulate the spiking activity of a Hodgkin-Huxley-like model, and how the tools built into our simulator make it easy to set up and integrate the model, and gain insight into the underlying biophysical mechanism. This simple model consists of a single-compartment neuron with three types of conductances  (Figure \ref{fig:figurehh}A). This heirarchical organization of the neuron is mirrored in the structure of the model: an object of type "compartment" is used to represent the cell body, and three objects of type "conductance" are used to model the three populations of ion channels (Figure \ref{fig:figurehh}B). These models of ion channels were obtained from \cite{liuModelNeuronActivityDependent1998} based on electrophysiological recordings from the lobster stomatogastric ganglion (\cite{turrigianoSelectiveRegulationCurrent1995}), and are part of the simulator. The syntax of constructing the neuron model (using "add" statements) preserves this herirarchical organization, and the code snippet in (Figure \ref{fig:figurehh}C) shows how this model can be created using just a few lines of code.

@@

"xolotl" can reproduce such a voltage clamp experiment $in silico$. Figure \ref{fig:figureclamp}B illustrates how a simple model with a single compartment and a single ion channel type can be set up and clamped to a desired voltage.  Integrating the model yields the current required to clamp the cell at that voltage. Here, we use a delayed-rectifier potassium conductance (\cite{liuModelNeuronActivityDependent1998}) and simulate a voltage-clamp experiment whose goal is to infer the activation function of this channel. First, the cell is clamped to a number of different voltages (Fig. \ref{fig:figureclamp}C) and the resultant clamping currents are measured by integrating the model (Fig. \ref{fig:figureclamp}D). The asymptotic clamping currents depend on the clamping voltage in a nonlinear manner (Fig. \ref{fig:figureclamp}E), since the open probabilities of the channel are functions of the membrane voltage. Assuming the reversal potential is known, Eq. \eqref{eq:2} can be used to solve for the total conductance of the channel as a function of the clamping voltage (Fig. \ref{fig:figureclamp}F). Finally, a sigmoid can be fit to the normalized conductance-voltage curves to obtain the activiation function of the ion channel population (Fig. \ref{fig:figureclamp}G-H). "xolotl" can therefore be used to describe graphically the theoretical underpinnings of ion channel characterization through voltage clamp and can serve as an effective pedagogical tool in computational neuroscience.

@@

Another intracellular mechanism of interest is a neuron's ability to regulate its electrical activity by controlling the spectrum of ion channels its expresses. Here, we will show "xolotl" can be used to model a recently proposed model of a homeostatic feedback system that controls the transcription rates of ion channels with the integral of an error signal derived from the intracellular Calcium concentration and some target Calcium (\cite{olearyCorrelationsIonChannel2013, olearyCellTypesNetwork2014}) (Figure \ref{fig:figuremechanism}H). Since this mechanism affects each ion channel population individually, object corresponding to this mechanism is added to each "condutance" object in the neuron (Figure \ref{fig:figuremechanism}I). Setting all maximal conductance densities to some low value and integrating the model shows that the intracellular Calcium levels rise over time and approach the target (Figure \ref{fig:figuremechanism}J), while all conductance densities increase and then remain bounded (Figure \ref{fig:figuremechanism}K). Examining the voltage dynamics of the cell reveal that it transitions from silence to truncated bursts of action potentials to periodic bursting as this mechanism regualtes the neuron's ion channel spectrum.

@@

"C" is statically-typed, with procedural syntax that provides low-level access to memory (\cite{kernighanProgrammingLanguage1978}, providing significant advantages for time-intensive computations. Unfortunately, automatic code-generation is limited by the supervening language. "MATLAB", for instance, cannot use pointers or pass by reference, which limits the efficiency of "C" code automatically generated from "MATLAB". Conversely, custom "C/C++" code provides significant increases in performance and memory conservation, but lacks the ease-of-use and flexibility of scripting languages.

@@

AH received funding from National Institute on Drug Abuse (NIDA) through the undergraduate training grant in computational neuroscience (1R90DA033463-01).

@@

\begin{figure}[!htb]
	\centering
	\includegraphics[width=1.0\linewidth]{gfx/figure_network}
	\caption{Simulating a network of conductance-based model neurons. (A) A three-compartment model of the pyloric network in the crustacean stomatogastric ganglion (cite Prinz \textit{et al.} 2004). Each neuron is modeled with a single compartment with 7-8 intrinsic conductances and 1-3 post-synaptic conductances. Synapses can be one of two types (Cholinergic (dashed lines) or Glutamatergic (solid lines)) and have different kinetics. (B) The code snippet shows how neurons can be created, wired together using synapses, and how the model can be integrated to return voltages and intracellular calcium levels in every compartment, and the state of every synapse. (C-E) Simulated voltage trace of a model network for the three compartments obtained from this simulation. (F) Time series activation variable of the Glutamatergic synapse between PY and LP (red connection in diagram) shows how the synapse becomes active every time the PY neuron spikes.}
	\label{fig:figurenetwork}
\end{figure}

>>>>

In this example, we demonstrate how to simulate the spiking activity of a Hodgkin-Huxley-like model, and how the tools built into our simulator make it easy to set up and integrate the model, and gain insight into the underlying biophysical mechanism. This simple model consists of a single-compartment neuron with three types of conductances  (Figure \ref{fig:figurehh}A). This hierarchical organization of the neuron is mirrored in the structure of the model: an object of type "compartment" is used to represent the cell body, and three objects of type "conductance" are used to model the three populations of ion channels (Figure \ref{fig:figurehh}B). These models of ion channels were obtained from \cite{liuModelNeuronActivityDependent1998} based on electrophysiological recordings from the lobster stomatogastric ganglion (\cite{turrigianoSelectiveRegulationCurrent1995}), and are part of the simulator. The syntax of constructing the neuron model (using "add" statements) preserves this hierarchical organization, and the code snippet in (Figure \ref{fig:figurehh}C) shows how this model can be created using just a few lines of code.

@@

"xolotl" can reproduce such a voltage clamp experiment $in silico$. Figure \ref{fig:figureclamp}B illustrates how a simple model with a single compartment and a single ion channel type can be set up and clamped to a desired voltage.  Integrating the model yields the current required to clamp the cell at that voltage. Here, we use a delayed-rectifier potassium conductance (\cite{liuModelNeuronActivityDependent1998}) and simulate a voltage-clamp experiment whose goal is to infer the activation function of this channel. First, the cell is clamped to a number of different voltages (Fig. \ref{fig:figureclamp}C) and the resultant clamping currents are measured by integrating the model (Fig. \ref{fig:figureclamp}D). The asymptotic clamping currents depend on the clamping voltage in a nonlinear manner (Fig. \ref{fig:figureclamp}E), since the open probabilities of the channel are functions of the membrane voltage. Assuming the reversal potential is known, Eq. \eqref{eq:2} can be used to solve for the total conductance of the channel as a function of the clamping voltage (Fig. \ref{fig:figureclamp}F). Finally, a sigmoid can be fit to the normalized conductance-voltage curves to obtain the activation function of the ion channel population (Fig. \ref{fig:figureclamp}G-H). "xolotl" can therefore be used to describe graphically the theoretical underpinnings of ion channel characterization through voltage clamp and can serve as an effective pedagogical tool in computational neuroscience.

@@

Another intracellular mechanism of interest is a neuron's ability to regulate its electrical activity by controlling the spectrum of ion channels its expresses. Here, we will show "xolotl" can be used to model a recently proposed model of a homeostatic feedback system that controls the transcription rates of ion channels with the integral of an error signal derived from the intracellular Calcium concentration and some target Calcium (\cite{olearyCorrelationsIonChannel2013, olearyCellTypesNetwork2014}) (Figure \ref{fig:figuremechanism}H). Since this mechanism affects each ion channel population individually, object corresponding to this mechanism is added to each "conductance" object in the neuron (Figure \ref{fig:figuremechanism}I). Setting all maximal conductance densities to some low value and integrating the model shows that the intracellular Calcium levels rise over time and approach the target (Figure \ref{fig:figuremechanism}J), while all conductance densities increase and then remain bounded (Figure \ref{fig:figuremechanism}K). Examining the voltage dynamics of the cell reveal that it transitions from silence to truncated bursts of action potentials to periodic bursting as this mechanism regualtes the neuron's ion channel spectrum.

@@

"C" is statically-typed, with procedural syntax that provides low-level access to memory (\cite{kernighanProgrammingLanguage1978}), providing significant advantages for time-intensive computations. Unfortunately, automatic code-generation is limited by the supervening language. "MATLAB", for instance, cannot use pointers or pass by reference, which limits the efficiency of "C" code automatically generated from "MATLAB". Conversely, custom "C/C++" code provides significant increases in performance and memory conservation, but lacks the ease-of-use and flexibility of scripting languages.

@@

AH received funding from the National Institute on Drug Abuse (NIDA) through the undergraduate training grant in computational neuroscience (1R90DA033463-01).

@@

\begin{figure}[!htb]
	\centering
	\includegraphics[width=1.0\linewidth]{gfx/figure_network}
	\caption{Simulating a network of conductance-based model neurons. (A) A three-compartment model of the pyloric network in the crustacean stomatogastric ganglion (\cite{prinzSimilarNetworkActivity2004}). Each neuron is modeled with a single compartment with 7-8 intrinsic conductances and 1-3 post-synaptic conductances. Synapses can be one of two types (Cholinergic (dashed lines) or Glutamatergic (solid lines)) and have different kinetics. (B) The code snippet shows how neurons can be created, wired together using synapses, and how the model can be integrated to return voltages and intracellular calcium levels in every compartment, and the state of every synapse. (C-E) Simulated voltage trace of a model network for the three compartments obtained from this simulation. (F) Time series activation variable of the Glutamatergic synapse between PY and LP (red connection in diagram) shows how the synapse becomes active every time the PY neuron spikes.}
	\label{fig:figurenetwork}
\end{figure}

### 2018-07-31 paper/supplementary/supplementary.tex
Add to supplementary information, fix some spelling

\begin{enumerate}
	\item EAGes, EAGmut, EAGwt (\cite{bronkRegulationEagCa22018})
	\item A-type (\cite{brookingsAutomaticParameterEstimation2014})
	\item MICurrent (\cite{caplanManyParameterSets2014a})
	\item CaT, H, Kd, NaV (\cite{dethierPositiveFeedbackCellular2015a})
	\item CaL, CaPump, KCa, Kd, NaV (\cite{drionDopaminePacemakerNeuron2017})
	\item Ap, At, Cal, KCa, KCaT, Proc (\cite{goldmanGlobalStructureRobustness2001a})
	\item Shaker (\cite{hardieNovelPotassiumChannels1991})
	\item Novel, Shab, Shaker (\cite{herasModulationVoltagedependentConductances2018})
	\item A-type, CaS, CaT, H, KCa, Kd, NaV (\cite{kisperskyIncreaseSodiumConductance2012a})
	\item NaP, NaT (\cite{linActivityDependentAlternativeSplicing2012})
	\item A-type, CaS, CaT, H, KCa, Kd, NaV (\cite{liuModelNeuronActivityDependent1998})
	\item \cite{liuModelNeuronActivityDependent1998} conductances with cached gating functions
	\item \cite{liuModelNeuronActivityDependent1998} conductances with forward Euler integration
	\item \cite{liuModelNeuronActivityDependent1998} conductances with temperature dependence
	\item Kd and NaV conductances from Int1, LG, and MCN1 cells (\cite{nadimFrequencyRegulationSlow1998})
	\item Cal (\cite{nadimSynapticDepressionCreates1999})
	\item Drosophila NaV (\cite{odowdVoltageclampAnalysisSodium1988})
	\item A-type, CaS, CaT, H, KCa, Kd, NaV (\cite{prinzAlternativeHandtuningConductancebased2003a})
	\item \cite{prinzAlternativeHandtuningConductancebased2003a} conductances with cached gating functions
	\item \cite{prinzAlternativeHandtuningConductancebased2003a} conductances with temperature dependence
	\item Kd and NaV from Int1 cells and modulatory input and transient LTS conductances from LG cells (\cite{rodriguezConvergentRhythmGeneration2013})
	\item proctolin modulatory input (\cite{sharpDynamicClampComputergenerated1993a})
	\item Cal and Kd from AB-PD, LP, and PY cells (\cite{soto-trevinoActivitydependentModificationInhibitory2001})
	\item A-type, CaT, and KCa from AB and PD cells, generic H, Kd, NaP, NaV, and proctolin conductances (\cite{soto-trevinoComputationalModelElectrically2005a})
	\item proctolin modulatory input (\cite{swensenModulatorsConvergentCellular2001a})
	\item A-type, CaS, CaT, H, KCa, Kd, NaP, NaV (\cite{turrigianoSelectiveRegulationCurrent1995a})
	\item Drosophila NaV (\cite{wicherNonsynapticIonChannels2001})
\end{enumerate}

>>>>

	\begin{enumerate}
		\item EAGes, EAGmut, EAGwt (\cite{bronkRegulationEagCa22018})
		\item A-type (\cite{brookingsAutomaticParameterEstimation2014})
		\item Proctolin modulatory input current (\cite{caplanManyParameterSets2014})
		\item CaT, H, Kd, NaV (\cite{dethierPositiveFeedbackCellular2015})
		\item CaL, CaPump, KCa, Kd, NaV (\cite{drionDopaminePacemakerNeuron2017})
		\item Ap, At, Cal, KCa, KCaT, Proc (\cite{goldmanGlobalStructureRobustness2001})
		\item Shaker (\cite{hardieNovelPotassiumChannels1991})
		\item Novel, Shab, Shaker (\cite{herasModulationVoltagedependentConductances2018})
		\item A-type, CaS, CaT, H, KCa, Kd, NaV (\cite{kisperskyIncreaseSodiumConductance2012})
		\item NaP, NaT (\cite{linActivityDependentAlternativeSplicing2012})
		\item A-type, CaS, CaT, H, KCa, Kd, NaV (\cite{liuModelNeuronActivityDependent1998})
		\item \cite{liuModelNeuronActivityDependent1998} conductances with cached gating functions
		\item \cite{liuModelNeuronActivityDependent1998} conductances with forward Euler integration
		\item \cite{liuModelNeuronActivityDependent1998} conductances with temperature dependence
		\item Kd and NaV conductances from Int1, LG, and MCN1 cells (\cite{nadimFrequencyRegulationSlow1998})
		\item Calcium current (\cite{nadimSynapticDepressionCreates1999})
		\item Drosophila NaV (\cite{odowdVoltageclampAnalysisSodium1988})
		\item A-type, CaS, CaT, H, KCa, Kd, NaV (\cite{prinzAlternativeHandtuningConductancebased2003})
		\item \cite{prinzAlternativeHandtuningConductancebased2003} conductances with cached gating functions
		\item \cite{prinzAlternativeHandtuningConductancebased2003} conductances with temperature dependence
		\item Kd and NaV from Int1 cells and modulatory input and transient LTS conductances from LG cells (\cite{rodriguezConvergentRhythmGeneration2013})
		\item proctolin modulatory input (\cite{sharpDynamicClampComputergenerated1993})
		\item Cal and Kd from AB-PD, LP, and PY cells (\cite{soto-trevinoActivitydependentModificationInhibitory2001})
		\item A-type, CaT, and KCa from AB and PD cells, generic H, Kd, NaP, NaV, and proctolin conductances (\cite{soto-trevinoComputationalModelElectrically2005})
		\item Proctolin modulatory input (\cite{swensenModulatorsConvergentCellular2001})
		\item A-type, CaS, CaT, H, KCa, Kd, NaP, NaV (\cite{turrigianoSelectiveRegulationCurrent1995})
		\item Drosophila NaV (\cite{wicherNonsynapticIonChannels2001})
	\end{enumerate}

\section{Parameters for Simulations}

	Two single-compartment models were simulated in this paper. The first, a Hodgkin-Huxley model with three conductances ("NaV", "Kd", and "Leak") and injected current was simulated for Figure 1 and 7. The second is a stomatogastric neuron model with eight conductances ("NaV", "CaT", "CaS", "ACurrent", "KCa", "Kd", "HCurrent", "Leak") and was simulated for Figure 3 and 7. Both models have dynamics as described in \cite{liuModelNeuronActivityDependent1998}.

	\subsection{Parameters for Hodgkin-Huxley model}

		\begin{center}
			\begin{tabular}{|l|l|}
				\hline 
				\textbf{Parameter Name} & \textbf{Value} \\ 
				\hline 
				Membrane capacitance ("HH.Cm") & 10 ${nF}/{mm^2}$ \\ 
				\hline 
				Surface area ("HH.A") & 0.01 $mm^2$ \\ 
				\hline 
				Fast sodium maximal conductance ("HH.NaV.gbar") & 1000 $\mu S/mm^2$ \\ 
				\hline 
				Delayed rectifier maximal conductance ("HH.Kd.gbar") & 300 $\mu S/mm^2$ \\ 
				\hline 
				Passive leak maximal conductance ("HH.Leak.gbar") & 1 $\mu S/mm^2$ \\ 
				\hline 
				Sodium reversal potential ("HH.NaV.E") & 50 $mV$ \\ 
				\hline 
				Potassium reversal potential ("HH.Kd.E") & -80 $mV$ \\ 
				\hline
				Leak reversal potential ("HH.Leak.E") & -50 $mV$ \\
				\hline
				Injected current ("I_ext") & 0.2 $nA$ \\
				\hline 
			\end{tabular} 
		\end{center}

	\subsection{Parameters for Stomatogastric Model}

		\begin{center}
			\begin{tabular}{|l|l|}
				\hline 
				\textbf{Parameter Name} & \textbf{Value} \\ 
				\hline 
				Membrane capacitance ("AB.Cm") & 10 ${nF}/{mm^2}$ \\ 
				\hline 
				Surface area ("AB.A") & 0.0628 $mm^2$ \\
				\hline
				Calcium buffering $\phi$ ("phi") & 90 $\mu M / nA$ \\
				\hline
				Calcium buffering shell volume ("vol") & 0.0628 $mm^3$ \\
				\hline 
				Fast sodium maximal conductance ("AB.NaV.gbar") & 00 $\mu S/mm^2$ \\ 
				\hline
				Transient calcium maximal conductance ("AB.CaT.gbar") & 00 $\mu S/mm^2$ \\ 
				\hline 
				Slow calcium maximal conductance ("AB.CaS.gbar") & 00 $\mu S/mm^2$ \\ 
				\hline 
				Transient potassium maximal conductance ("AB.ACurrent.gbar") & 00 $\mu S/mm^2$ \\ 
				\hline 
				Calcium-gated potassium maximal conductance ("AB.KCa.gbar") & 00 $\mu S/mm^2$ \\ 
				\hline 
				Delayed rectifier maximal conductance ("AB.Kd.gbar") & 00 $\mu S/mm^2$ \\ 
				\hline
				Hyperpolarization-activated maximal conductance ("AB.HCurrent.gbar") & 00 $\mu S/mm^2$ \\ 
				\hline 
				Passive leak maximal conductance ("AB.Leak.gbar") & 00 $\mu S/mm^2$ \\ 
				\hline 
				Sodium reversal potential ("AB.NaV.E") & 30 $mV$ \\ 
				\hline 
				Potassium reversal potential ("AB.Kd.E") & -80 $mV$ \\ 
				\hline
				Hyperpolarization-activated reversal potential ("AB.HCurrent.E") & -20 $mV$ \\
				\hline
				Leak reversal potential ("AB.Leak.E") & -50 $mV$ \\
				\hline
			\end{tabular} 
		\end{center}

	\subsection{Parameters for Network Model}
		The network model displayed in Figure 4 is described in \cite{prinzAlternativeHandtuningConductancebased2003, prinzSimilarNetworkActivity2004}. It is comprised of three compartments "ABPD", "LP", and "PY" with the same dynamics but differing parameters and synaptic inputs.

		\begin{center}
			\begin{tabular}{|l|c|c|c|c|}
				\hline 
				\textbf{Parameter Name} & "ABPD" & "LP" & "PY" & \textbf{Units} \\ 
				\hline 
				Fast sodium maximal conductance & 1000 & 1000 & 1000 & $\mu S/mm^2$ \\ 
				\hline 
				Transient calcium maximal conductance & 25 & 0 & 24 & $\mu S/mm^2$ \\  
				\hline 
				Slow calcium maximal conductance & 60 & 40 & 20 & $\mu S/mm^2$ \\  
				\hline 
				Transient calcium maximal conductance & 500 & 200 & 500 & $\mu S/mm^2$ \\  
				\hline 
				Calcium-gated potassium maximal conductance & 50 & 0 & 0 & $\mu S/mm^2$ \\  
				\hline 
				Delayed rectifier maximal conductance & 1000 & 250 & 1250 & $\mu S/mm^2$ \\  
				\hline 
				Hyperpolarization-activated maximal conductance & 0.1 & 0.5 & 0.5 & $\mu S/mm^2$ \\  
				\hline 
				Leak maximal conductance & 0 & 0.3 & 0.1 & $\mu S/mm^2$ \\  
				\hline 
			\end{tabular}
		\end{center}

		There are seven synapses in the model of two types: glutamatergic ("Glut") and cholinergic ("Chol") with differing dynamics. The synapse produces a current in the postsynaptic compartment based on the membrane potential in the presynaptic compartment.

		\begin{center}
			\begin{tabular}{c|c|c|c|}
				\hline
				\textbf{Presynaptic} & \textbf{Postsynaptic} & \textbf{Type} & \textbf{Maximal Conductance} \\
				\hline

			\end{tabular}
		\end{center}
